### 2019-10-29 Paper_shortened_pdf.Rmd
Reviewer 2 comment 8

Our target sample consisted of 1884 native `r masked("Australian")`, non-minority, intact, biological families. At the first wave, children were approximately 6 years old, $M = 6.28$, $SD = .45$, and 12 at the final wave ($M = 12.42$, $SD = .49$); on agerage, mothers were 34.89 ($SD = 4.71$), and fathers 36.94 ($SD = 5.48$). Most families were dual-earner families, based on income from wages or enterprise ($N = 1067$); the remainder were single-earner families($N = 669$ father-only, $N = 41$ mother-only). The modal annual income for fathers was between \$26,000-\$51,999 (N = 663), and for mothers, \$25,999 or less per year (N = 1152).

@@

Historically, developmental theory has focused on how parents cultivate their children's development. Nowadays, it is commonly acknowledged that influences are reciprocal. However, commonly used developmental methods have been criticized for misrepresenting the direction of effects when stable between-family differences are not accounted for [@hamakerCritiqueCrosslaggedPanel2015]. One recent study found that, in adolescence, child effects substantially outnumbered parenting effects [@vanlissaRoleFathersMothers2018a]. The authors speculated that this might be due to the greater equality of parent-child relationships in adolescence, compared to earlier childhood. The present study casts doubt on this interpretation, as we found that child effects outweighed parenting effects, even at a young age. The prominence of child effects across two large longitudinal studies, with samples from different continents and at different age ranges, invites the question: Are parents really shaping children's development, or are they, instead, adapting to it? An extreme interpretation of our findings might be that parents "set the stage" for children's development, through their stable differences from other parents, and from there on, are merely responding to children's changing developmental needs. This extreme view can be nuanced by taking into consideration that not all causal effects are evident from the interplay between within-individual fluctuations. For example, although the genetic makeup that is passed from parents to children is stable, the expression of some genes emerges over time. Consequently, stable parent characteristics can cause change in children. Future research might examine alternative theoretical conceptualizations of parental influences in child development. Another point to consider is that parent effects on children might play out on a different time scale than was captured by these macro-longitudinal studies. If parents shape their children through specific behaviors in dyadic interactions, then we might fail to capture parenting effects when using annual reflective self-reports.

>>>>

Our target sample consisted of 1884 native `r masked("Australian")`, non-minority, intact, biological families. At the first wave, children were approximately 6 years old, $M = 6.28$, $SD = .45$, and 12 at the final wave ($M = 12.42$, $SD = .49$); on agerage, mothers were 34.89 ($SD = 4.71$), and fathers 36.94 ($SD = 5.48$). Most families were dual-earner families, based on income from wages or enterprise ($N = 1067$); the remainder were single-earner families($N = 669$ father-only, $N = 41$ mother-only). The modal annual income for fathers was between \$26,000-\$51,999 ($N = 663$ fathers reported this value), and for mothers, \$25,999 or less per year ($N = 1152$ reported this value).

@@

Historically, developmental theory has focused on how parents cultivate their children's development. Nowadays, it is commonly acknowledged that influences are reciprocal. However, commonly used developmental methods have been criticized for misrepresenting the direction of effects when stable between-family differences are not accounted for [@hamakerCritiqueCrosslaggedPanel2015]. One recent study with a similar design to the present study found that child effects substantially outnumbered parenting effects, and speculated that this might be due to the greater equality of parent-adolescent relationships, compared to earlier childhood [@vanlissaRoleFathersMothers2018a]. The present study casts doubt on this speculation, as we found that child effects outweighed parenting effects, even at a young age. The fact that both the present study and this previous work found a preponderance of child effects, despite the different age ranges of the samples, invites the question: Are parents really shaping children's development, or are they, instead, adapting to it? An extreme interpretation of our findings might be that parents "set the stage" for children's development, through their stable differences from other parents, and from there on, are merely responding to children's changing developmental needs. This extreme view can be nuanced by taking into consideration that not all causal effects are evident from the interplay between within-individual fluctuations. For example, although the genetic makeup that is passed from parents to children is stable, the expression of some genes emerges over time. Consequently, stable parent characteristics can cause change in children. Future research might examine alternative theoretical conceptualizations of parental influences in child development. Another point to consider is that parent effects on children might play out on a different time scale than was captured by these macro-longitudinal studies. If parents shape their children through specific behaviors in dyadic interactions, then we might fail to capture parenting effects when using annual reflective self-reports.
